Fully Automatic Loop-free Program Synthesis by Bounded Model Checking

### 2014-01-21 paper/paper.tex
Add details of program encoding

\usepackage{listings}
\usepackage{graphicx}
\usepackage{xspace}
\usepackage{pgf}
\usepackage[noend]{algpseudocode}
\usepackage{algorithm}
\usepackage{amsfonts}
\usepackage{pifont}

@@

@@

Our encoding is the following:
%
\begin{itemize}
 \item $I$ is the set of $N$-tuples of 32-bit bitvectors.  This is written in \newC as the type \verb|int[N]|.
 \item $O$ is the set of $M$-tuples of 32-bit bitvectors, which is written in \newC as the type \verb|int[M]|.
 \item $\sigma$ is a pure function with type $I \times O \rightarrow \mathrm{Bool}$.  The \newC signature of this function is
 \verb|int check(int in[N], int out[M])|. This function is the only component supplied
 by the user.
 \item $P$ is written in a simple RISC-like language $\mathcal{L}$, described in Fig.~\ref{fig:l-language}.  Programs in $\mathcal{L}$
 have type $I \rightarrow O$ and
 are represented in \newC as objects of type \verb|prog_t|.
 \item We supply an interpreter for $\mathcal{L}$ which is written in \newC.  The type of
 this interpreter is $(I \rightarrow O) \times I \rightarrow O$ and the \newC signature is \\
 \verb|void exec(prog_t p, int in[N], int out[M])|.  Here, \verb|out| is an output parameter.
\end{itemize}

@@

We must now express the {\sc synth} and {\sc verif} formulae as safety properties
of \newC programs, which is given in Fig.~\ref{fig:c-synthverif}.

@@

\section{Parametrisation of the Program Space}

@@

>>>>

\usepackage{listings}
\usepackage{graphicx}
\usepackage{xspace}
\usepackage{pgf}
\usepackage[noend]{algpseudocode}
\usepackage{algorithm}
\usepackage{amsfonts}
\usepackage{amsmath}
%\usepackage{amsthm}
\usepackage{pifont}

@@

\newtheorem{defn}{Definition}

@@

Our encoding is the following:
%
\begin{itemize}
 \item $I$ is the set of $N$-tuples of 32-bit bitvectors.  This is written in \newC as the type \verb|int[N]|.
 \item $O$ is the set of $M$-tuples of 32-bit bitvectors, which is written in \newC as the type \verb|int[M]|.
 \item $\sigma$ is a pure function with type $I \times O \rightarrow \mathrm{Bool}$.  The \newC signature of this function is
 \verb|int check(int in[N], int out[M])|. This function is the only component supplied
 by the user.
 \item $P$ is written in a simple RISC-like language $\mathcal{L}$, whose syntax is given in Fig.~\ref{fig:l-language}.  Programs in $\mathcal{L}$
 have type $I \rightarrow O$ and
 are represented in \newC as objects of type \verb|prog_t|, shown in Fig.~\ref{fig:c-l-encoding}.
 \item We supply an interpreter for $\mathcal{L}$ which is written in \newC.  The type of
 this interpreter is $(I \rightarrow O) \times I \rightarrow O$ and the \newC signature is \\
 \verb|void exec(prog_t p, int in[N], int out[M])|.  Here, \verb|out| is an output parameter.
\end{itemize}

@@

The exact details of how we encoding an $\mathcal{L}$-program are given in Sec.~\ref{sec:encode-l}.
We must now express the {\sc synth} and {\sc verif} formulae as safety properties
of \newC programs, which is given in Fig.~\ref{fig:c-synthverif}.

@@

\section{Parameterising the Program Space}

@@

\subsection{Encoding an $\mathcal{L}$-Program in \newC}
\label{sec:encode-l}

\begin{figure}
\begin{lstlisting}[language=c++,mathescape]
typedef $\mathcal{BV}(4)$ op_t;                  // An opcode
typedef $\mathcal{BV}(w)$ word_t;                // An $\mathcal{L}$-word
typedef $\mathcal{BV}(\log_2 \lceil c+l+a \rceil)$ param_t;  // An operand

struct prog_t {
  op_t ops[$l$];          // The opcodes
  param_t params[$l$*2];  // The operands
  word_t consts[$c$];     // The program constants
}
\end{lstlisting}

 \caption{The \newC structure we use to encode an $\mathcal{L}$ program
  \label{fig:c-l-encoding}}
\end{figure}

The exact \newC encoding of an $\mathcal{L}$ program is shown in Fig.~\ref{fig:c-l-encoding}.
The \verb|prog_t| structure encodes a complete program, which is a sequence of instructions.
The parameter $a$ is the number of arguments the program takes.
The $i$th instruction has opcode \verb|ops[i]|, left operand \verb|params[i*2]| and
right operand \verb|params[i*2 + 1]|.  An operand refers to either a program constant,
a program argument or the result of a previous instruction.  The value of an operand
is determined at runtime as follows:

\[
 val(x) = \begin{cases}
           x < a & \text{the } x^{\text{th}} \text{ program argument} \\
           a \leq x < a+c & \mathtt{consts[} x-a \mathtt{]} \\
           x \geq a + c & \text{the result of the } (x - a - c)^{\text{th}} \text{ instruction}
          \end{cases}
\]

\begin{defn}
A program is well formed if
no operand refers to the result of an instruction that has not been computed yet.
\label{defn:well-formed}
\end{defn}

We add a well-formedness constraint of the form \verb|params[i] <= (a + c + 2*i)|
for each instruction.  It should be noted that this requires a linear number of well-formedness
constraints.  If all of these constraints are satisfied, the program is well-formed in the sense
of Defn.~\ref{defn:well-formed}.
